---
title: "CodeBook"
author: "Sarvesh Amberkar"
date: "01/07/2020"
output: pdf_document
---
## Download the dataset:


Dataset downloaded from *https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip* and extracted under the folder called **UCI HAR Dataset**


```{r reading,include=FALSE}
library(dplyr)
library(data.table)

features <- read.table("features.txt", col.names = c("n","functions"))
activity<-read.table("activity_labels.txt",col.names = c("n","labels"))
x_test <- read.table("X_test.txt", col.names = features$functions)
y_test <- read.table("y_test.txt", col.names = "code")
subject_train <- read.table("subject_train.txt", col.names = "subject")
subject_test <- read.table("subject_test.txt", col.names = "subject")
x_train <- read.table("X_train.txt", col.names = features$functions)
y_train <- read.table("y_train.txt", col.names = "code")
```


## 1. Assigning tables to variable:


### a) features<-features.txt 

*561 rows, 2 columns*

The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ. These time domain signals (prefix 't' to denote time) were captured at a constant rate of 50 Hz


### b) Activities <-activity_labels.txt 

*6 rows, 2 columns*

List of activities performed when the corresponding measurements were taken.
labels are  WALKING , WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING.


### c) x_test <- test/X_test.txt

*2947 rows, 561 columns*

recorded features of x_test


### d) y_test <- test/y_test.txt :

*2947 rows, 1 columns*

contains test data of activities code labels


### e) subject_train <- test/subject_train.txt : 

*7352 rows, 1 column*

contains train data recorded using sensor.


### f) y_train <- test/y_train.txt :

*7352 rows, 1 columns*

contains train data of activities code labels


## 2. Merges the training and the test sets to create one data set

merging train and test data using rbind.
```{r merge,include= FALSE}
X<-rbind(x_train,x_test)
Y<-rbind(y_train,y_test)
Subject<-rbind(subject_train,subject_test)
Merged_Data <- cbind(Subject, Y, X)

```


## 3. Extracts only the measurements on the mean and standard deviation for each measurement

TidyData (10299 rows, 88 columns) is created by subsetting Merged_Data
```{r extract,include=FALSE}
a<-Merged_Data[,grep("(*mean*)|(*std*)",tolower(colnames(Merged_Data)))]
a<-cbind(Merged_Data$code,a)
colnames(a)[1]<-"code" 
a<-cbind(Merged_Data$subject,a)
colnames(a)[1]<-"subject" 

```



## 4. Uses descriptive activity names to name the activities in the data set

Entire numbers in code column of the TidyData replaced with corresponding activity 

```{r labels,include=FALSE}
a$code<-as.factor(a$code)
levels(a$code)<-activity$labels
```

## 5. Appropriately labels the data set with descriptive variable names

short form of some name replaced by the full form example (t in start denotes time),(Gyro with Gyroscope)
```{r namess,include=FALSE}
names(a)[2] = "activity"
names(a)<-gsub("Acc", "Accelerometer", names(a))
names(a)<-gsub("^t", "Time", names(a))
names(a)<-gsub("^f", "Frequency", names(a))

names(a)<-gsub("BodyBody", "Body", names(a))


```


## 6. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject
```{r newtidy,include=FALSE}

tidy_data_mean<-tbl_df(a) %>% group_by(activity,subject) %>% summarize_all(mean)
```

Taking the means of each variable for each activity and each subject, after groupped by subject and activity.

## summary of the tidy data set

```{r summarizing,echo=FALSE}
summary(a)

```






